Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-08-09 Report/simulation_distributed.tex
updated figures

The performance of the \ac{MSE}-\ac{KKT} scheme discussed in Section \ref{sec-4.3} is presented in Fig. \ref{fig-d-3}. It compares the \ac{JSFRA} scheme with \me{\ell_1} and \me{\ell_2} norm cases in addition to the \ac{Q-WSRM} scheme. We compare the performance of the \ac{MSE}-\ac{KKT} scheme for the exponents \me{q=1} and \me{q=2}. The performance of the \me{\ell_1} \ac{JSFRA} scheme performs the best over other algorithms in minimizing the total number of queued bits after the current allocation due to its greedy approach. The inclusion of the additional second order rate term in \eqref{pf-1}, the \me{\ell_2} \ac{JSFRA} scheme achieves different performance than the \ac{Q-WSRM} scheme as shown in Fig. \ref{fig-d-3}.

It can be seen from Fig. \ref{fig-d-3.2}, that the sum rate of the \me{q=1} \ac{KKT} approach outperforms other schemes in spite of Fig. \ref{fig-d-3.1} showing inferior queue minimizing performance. This due to the lack of the rate bounding constraint \eqref{kkt-mse-4.2}, which is simply given by \me{\sigma_{l,k,n}^{(i)} = \frac{a_k}{\log(2)}}, providing no control for the excess allocations. For \me{q=2}, when the allocated rate exceeds the available queued bits, \me{\sigma_{l,k,n}^{(i)}} will be negative, thereby reducing \eqref{kkt-mse-4.1} as \me{\alpha_{l,k,n}^{(i)} < \alpha_{l,k,n}^{(i-1)}} to reduce the priority for the corresponding user in the next iteration. The performance of the closed form solution using the \ac{KKT} solution performs similar to the centralized schemes with \me{q=2} by limiting the rates beyond the available queued packets.
\begin{figure*}
\centering
\subfloat[][Queue deviation]{
\includegraphics[width=0.48\textwidth]{fig-9}
\label{fig-d-3.1}}
\hfill
\subfloat[][Sum rate performance]{
\includegraphics[width=0.48\textwidth]{fig-10}
\label{fig-d-3.2}}
\caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace} model}
\label{fig-d-3}
\end{figure*}

>>>>

The performance of the \ac{MSE}-\ac{KKT} scheme discussed in Section \ref{sec-4.3} is presented in Fig. \ref{fig-d-3}. It compares the \ac{JSFRA} scheme with \me{\ell_1} and \me{\ell_2} norm cases in addition to the \ac{Q-WSRM} scheme. We compare the performance of the \ac{MSE}-\ac{KKT} scheme for the exponents \me{q=1} and \me{q=2}. The performance of the \me{\ell_1} \ac{JSFRA} scheme performs the best over other algorithms in minimizing the total number of queued bits after the current allocation due to the greedy approach of the objective. The inclusion of the additional second order rate term in \eqref{pf-1}, the \me{\ell_2} \ac{JSFRA} scheme achieves different performance than the \ac{Q-WSRM} scheme as shown in Fig. \ref{fig-d-3}.

It can be seen from Fig. \ref{fig-d-3.2}, that the sum rate of the \me{q=1} \ac{KKT} approach outperforms other schemes in spite of Fig. \ref{fig-d-3.1} showing inferior queue minimizing performance. This due to the lack of the rate bounding constraint \eqref{kkt-mse-4.2}, which is simply given by \me{\sigma_{l,k,n}^{(i)} = \frac{a_k}{\log(2)}}, providing no control for the excess allocations. For \me{q=2}, when the allocated rate exceeds the available queued bits, \me{\sigma_{l,k,n}^{(i)}} will be negative, thereby reducing \eqref{kkt-mse-4.1} as \me{\alpha_{l,k,n}^{(i)} < \alpha_{l,k,n}^{(i-1)}} to reduce the priority for the corresponding user in the next iteration. The performance of the closed form solution using the \ac{KKT} solution performs similar to the centralized schemes with \me{q=2} by limiting the rates beyond the available queued packets.
\begin{figure*}
\centering
\subfloat[][Queue deviation]{
\includegraphics[width=0.48\textwidth]{fig-9-2}
\label{fig-d-3.1}}
\hfill
\subfloat[][Sum rate performance]{
\includegraphics[width=0.48\textwidth]{fig-10}
\label{fig-d-3.2}}
\caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace} model}
\label{fig-d-3}
\end{figure*}

### 2014-08-09 Report/simulation_results.tex
stfloats
updated figures

This section discusses the performance of the centralized algorithms discussed in Section \ref{sec-3} for a few system configurations. To begin with, we consider a single cell \ac{SISO} model operating at \me{10} dB \ac{SNR} with \me{K = 3} users sharing \me{N = 3} sub-channel resources. The number of packets waiting at the transmitter for each user is given by \me{Q_k = 5,8} and \me{6} bits respectively. With this configuration, the resources allocated for the users on each sub-channel by various algorithms are listed in Table \ref{tbl-1}.

Table \ref{tbl-1} lists the sub-channel wise channel seen by the users followed by the assigned rates over each sub-channel by three different algorithms, \ac{Q-WSRM} allocation, \ac{JSFRA} scheme and the band-wise \ac{Q-WSRM} scheme using \ac{WMMSE} precoder design proposed in \cite{wmmse_shi}. The performance metric used for the comparison of different algorithms is the total number of backlogged bits left over at each slot after the allocation, which is denoted by \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. It can be seen from Table \ref{tbl-1}, that the \ac{Q-WSRM} scheme provides more priority to the third user with \me{Q_{3} = 6} bits compared to the first user with \me{Q_{1} = 5} bits while allocating the first sub-channel. Int contrast to the \ac{Q-WSRM} scheme, the \ac{JSFRA} scheme assigns the first user on the first sub-channel thereby reducing the total number of backlogged packets waiting at the transmitter.
\begin{table*}
\centering
\renewcommand{\arraystretch}{1.25} \scriptsize
\begin{tabular}{|*{14}{c|}}
\hline
\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRM approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(\emph{backpreassure} algorithm)} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
\cline{3-14}
 && Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} \\
\hline
\me{1} & \me{5} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.91} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.46}  &  \me{3.54} & \me{0} &  \me{4.36}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
\me{3} & \me{6} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{5.72} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{5.79} & \me{5.81} &  \me{0}  &  \me{0} \\
\hline
\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{5.28} bits} & \multicolumn{3}{c|}{\me{3.91} bits} & \multicolumn{3}{c|}{\me{5.28} bits} \\
\hline
\end{tabular}
\caption{Sub channel wise allocation for a scheduling instant}
\label{tbl-1}
\end{table*}

In order to understand the behavior in a \ac{MIMO} framework, we consider a system with \me{N=3} sub-channels and \me{N_B = 3} \acp{BS}, each equipped with \me{N_T = 4} transmit antennas operating at \me{10}dB \ac{SNR}, serving \me{|\mc{U}_b| = 3} users each. The users are located with the maximum interference seen from the neighboring \acp{BS} is limited to \me{-6} dB, thereby simulating a realistic scenario. Fig. \ref{fig-1} shows the performance of the centralized schemes for a single receive antenna system. It compares the total number of \ac{SCA} updates required by the \ac{JSFRA}, \ac{SRA} and the \ac{Q-WSRM} schemes to perform the optimal allocations to minimize the total number of backlogged packets.
\begin{figure*}
\centering
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=1.0\textwidth]{fig-1-1}
\caption{\me{4 \times 1} system}
\label{fig-1}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=1.0\textwidth]{fig-2-1}
\caption{\me{4 \times 2} system}
\label{fig-2}
\end{subfigure}
\caption{Convergence plot for \me{\lbrace N,N_B,K \rbrace = \lbrace 3,3,9 \rbrace} model}
\end{figure*}

In the sub-channel wise allocations, where the total transmit power is shared equally among the sub-channels, the precoders are designed for each sub-channels independently. In this approach, the precoders are coupled via the number of queued bits which are updated using \eqref{eqn-weight} before designing the precoders for the sub-channels. At each \ac{SCA} points, the number of queued bits are reduced significantly with the introduction of a new sub-channel since the algorithm starts with a random initial point before it converges to an optimal\footnote{due to the nonconvex nature of the original problem} precoders. The total number of backlogged bits at each \ac{SCA} update instant are plotted in Fig. \ref{fig-1} for the discussed centralized schemes and the convergence point are marked with the data tips. Fig. \ref{fig-2} compares the performance of the centralized algorithms for the \me{N_R = 2} receive antennas.
\begin{table}
\centering
\renewcommand{\arraystretch}{1.25} \scriptsize
\begin{tabular}{|c|*{8}{c}|c|}
\hline
\me{q} & \multicolumn{8}{c|}{user indices} & \me{\chi} \\
\hline
\me{1} & 15.00 & 3.95 & 5.26 & 8.95 & 7.03 & 11.90 & 12.00 & 9.73 & 25.15 \\
\me{2} & 11.23 & 3.93 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.90 & 27.77 \\
\me{\infty} & 11.41 & 4.41 & 10.41 & 10.41 & 10.41 & 8.41 &  8.41 &  6.41 & 28.68 \\
\hline
\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\    
\cline{1-9}
\end{tabular}
\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
\label{tbl-3}
\end{table}

The behavior of the \ac{JSFRA} algorithm for different exponents \me{q} are outlined in the Table \ref{tbl-3} for the users located at the cell-edge of the system employing \me{N_T = 4} transmit antennas. The configuration is mentioned in the caption of Table \ref{tbl-3} along with the number of queued bits for each user. It is evident that the algorithm minimizes the queued bits for the \me{\ell_1} norm compared to the \me{\ell_2} norm, which is shown in the column displaying the total number of left over packets \me{\chi} in bits. The \me{\ell_{\infty}} norm provides fair allocation of the resources by making the left over packets \me{\chi_k = 3.58} bits. The \me{\ell_{\infty}} norm provides the fair allocation by making the queued deviation equal for all the users after the current allocation irrespective of their channel gains.

>>>>

This section discusses the performance of the centralized algorithms discussed in Section \ref{sec-3} for a few system configurations. To begin with, we consider a single cell \ac{SISO} model operating at \me{10} dB \ac{SNR} with \me{K = 3} users sharing \me{N = 3} sub-channel resources. The number of packets waiting at the transmitter for each user is given by \me{Q_k = 5,8} and \me{6} bits respectively. With this configuration, the resources allocated for the users on each sub-channel by various algorithms are shown in Fig. \ref{tbl-1}.

Fig. \ref{tbl-1} plots the sub-channel wise channel seen by the users followed by the assigned rates over each sub-channel by three different algorithms, \ac{Q-WSRM} allocation, \ac{JSFRA} scheme and the band-wise \ac{Q-WSRM} scheme using \ac{WMMSE} precoder design proposed in \cite{wmmse_shi}. The performance metric used for the comparison of different algorithms is the total number of backlogged bits left over at each slot after the allocation, which is denoted by \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. It can be seen from Fig. \ref{tbl-1}, that the \ac{Q-WSRM} scheme provides more priority to the third user with \me{Q_{3} = 6} bits compared to the first user with \me{Q_{1} = 5} bits while allocating the first sub-channel. Int contrast to the \ac{Q-WSRM} scheme, the \ac{JSFRA} scheme assigns the first user on the first sub-channel thereby reducing the total number of backlogged packets waiting at the transmitter.
%\begin{table*}
%\centering
%\renewcommand{\arraystretch}{1.25} \scriptsize
%\begin{tabular}{|*{14}{c|}}
%\hline
%\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRM approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
%\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(\emph{backpreassure} algorithm)} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
%\cline{3-14}
% && Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} & Sch-\me{1} & Sch-\me{2} & Sch-\me{3} \\
%\hline
%\me{1} & \me{5} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.91} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
%\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.46}  &  \me{3.54} & \me{0} &  \me{4.36}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
%\me{3} & \me{6} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{5.72} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{5.79} & \me{5.81} &  \me{0}  &  \me{0} \\
%\hline
%\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{5.28} bits} & \multicolumn{3}{c|}{\me{3.91} bits} & \multicolumn{3}{c|}{\me{5.28} bits} \\
%\hline
%\end{tabular}
%\caption{Sub channel wise allocation for a scheduling instant}
%\label{tbl-1}
%\end{table*}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{tbl-1}
\caption{Allocation for a \ac{SISO} Model comparing different Algorithms}
\label{tbl-1}
\end{figure}

In order to understand the behavior in a \ac{MIMO} framework, we consider a system with \me{N=3} sub-channels and \me{N_B = 3} \acp{BS}, each equipped with \me{N_T = 4} transmit antennas operating at \me{10}dB \ac{SNR}, serving \me{|\mc{U}_b| = 3} users each. The users are located with the maximum interference seen from the neighboring \acp{BS} is limited to \me{-6} dB, thereby simulating a realistic scenario. Fig. \ref{fig-1} shows the performance of the centralized schemes for a single receive antenna system. It compares the total number of \ac{SCA} updates required by the \ac{JSFRA}, \ac{SRA} and the \ac{Q-WSRM} schemes to perform the optimal allocations to minimize the total number of backlogged packets.
\begin{figure*}
\centering
\subfloat[][\me{N_T \times N_R = 4 \times 1} system]{
\includegraphics[width=0.48\textwidth]{fig-1-2}
\label{fig-1}}
\hfill
\subfloat[][\me{N_T \times N_R = 4 \times 2} system]{
\includegraphics[width=0.48\textwidth]{fig-2-1}
\label{fig-2}}
\caption{Number of backlogged packets at the \ac{SCA} update points for \me{\lbrace N,N_B,K \rbrace = \lbrace 3,3,9\rbrace} model}
\label{fig-a}
\end{figure*}

In the sub-channel wise allocations, where the total transmit power is shared equally among the sub-channels, the precoders are designed for each sub-channels independently. In this approach, the precoders are coupled via the number of queued bits which are updated using \eqref{eqn-weight} before designing the precoders for the sub-channels. At each \ac{SCA} points, the number of queued bits are reduced significantly with the introduction of a new sub-channel since the algorithm starts with a random initial point before it converges to an optimal\footnote{due to the nonconvex nature of the original problem} precoders. The total number of backlogged bits at each \ac{SCA} update instant are plotted in Fig. \ref{fig-1} for the discussed centralized schemes and the convergence point are marked with the data tips. Fig. \ref{fig-2} compares the performance of the centralized algorithms for the \me{N_R = 2} receive antennas.
%\begin{table}
%\centering
%\renewcommand{\arraystretch}{1.25} \scriptsize
%\begin{tabular}{|c|*{8}{c}|c|}
%\hline
%\me{q} & \multicolumn{8}{c|}{user indices} & \me{\chi} \\
%\hline
%\me{1} & 15.00 & 3.95 & 5.26 & 8.95 & 7.03 & 11.90 & 12.00 & 9.73 & 25.15 \\
%\me{2} & 11.23 & 3.93 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.90 & 27.77 \\
%\me{\infty} & 11.41 & 4.41 & 10.41 & 10.41 & 10.41 & 8.41 &  8.41 &  6.41 & 28.68 \\
%\hline
%\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\
%\cline{1-9}
%\end{tabular}
%\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
%\label{tbl-3}
%\end{table}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{tbl-2}
\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
\label{tbl-3}
\end{figure}

The behavior of the \ac{JSFRA} algorithm for different exponents \me{q} are outlined in the Fig. \ref{tbl-3} for the users located at the cell-edge of the system employing \me{N_T = 4} transmit antennas. The configuration is mentioned in the caption of Fig. \ref{tbl-3} along with the number of queued bits for each user. It is evident that the algorithm minimizes the queued bits for the \me{\ell_1} norm compared to the \me{\ell_2} norm, which is shown in the column displaying the total number of left over packets \me{\chi} in bits. The \me{\ell_{\infty}} norm provides fair allocation of the resources by making the left over packets \me{\chi_k = 3.58} bits. The \me{\ell_{\infty}} norm provides the fair allocation by making the queued deviation equal for all the users after the current allocation irrespective of their channel gains.
